---
title: "Title"
date: "`r conr::format_date()`"
author: "Connor Ballinger"
knit: conr::write_and_date
output: conr::format_html
---

```{r, warning=FALSE, message=FALSE}
# knitr::opts_chunk$set()

library(here)
library(tidyverse)
library(knitr)
library(hesim)
library(data.table)

registerS3method("knit_print", "data.frame", conr::knit_print.data.frame)

# theme_set(theme_bw() + theme(legend.position = "bottom"))

```

# Overview

-   Epithelial OC (90% of OC is used).

-   3-state Markov (stable, progressed, death), stratified by the 4 stages. 

-   10-year horizon, societal perspective.

-   “The cycle length used was 3 weeks (21 days), which is the length of a chemotherapy cycle. Patients entered the model after they were diagnosed with EOC in the stable state. A cycle after, they can either remain stable, become worse, and move to a post-progression state, or die. Those who progress remain in a post-progression state until they die.” 

-   Used Heintz 2006 for survival. Heintz only provided 5-year survival data – Delgado extrapolated based on exponential distribution? ie time-constant transition probabilities.

-   Recall there may be no treatment effects applied - may just use Olaparib survival curves directly in an appropriate sub-population.

# Following DARTH Framework

## Structure

```{r}
n_cycle_length <- 3 / 52 # 3 weeks
n_cycles <- 10 * 52 / 3 # 10 years
n_discount_rate <- 0.03

v_names_states <- c("Stable", "Progressed", "Dead")
v_names_stages <- c("I", "II", "III", "IV")
```

## Transitions

-   Formulas in Delgado suggest probabilities vary over time, but they have assumed exponential so I'm not sure.

```{r}
n_bgr_mortality <- 0.001 # Place-holder

v_median_pfs <- c(18.33, 6.25, 2.00, 1.60)
v_median_os <- c(19.50, 7.50, 3.20, 1.90)

v_lambda_pfs <- - log(0.5) / v_median_pfs
v_3r_pfs <- v_lambda_pfs / (52/3) # change to 3-weekly rate
v_p_pp <- 1 - exp(-v_3r_pfs)

v_lambda_os <- - log(0.5) / (v_median_os - v_median_pfs)
v_3r_os <- v_lambda_os / (52/3) # change to 3-weekly rate
v_p_d <- 1 - exp(-v_3r_os)


m_trans <- list(v_p_pp, v_p_d) |> 
  pmap(\(x, y) matrix(data = c(
    (1 - x - n_bgr_mortality), x, n_bgr_mortality,
    0, (1 - y - n_bgr_mortality), y + n_bgr_mortality,
    0, 0, 1
  ), byrow = TRUE, nrow = 3, 
  dimnames = list(c("S", "P", "D"), c("S", "P", "D"))))

names(m_trans) <- v_names_stages
m_trans
```

```{r}
v_init_stages <- c(1155, 195, 1116, 681)
v_init_states <- c(1, 0, 0) # all patients start in stable state

m_trace <- v_names_stages |> 
  map(\(x) matrix(as.numeric(NA), nrow = (n_cycles + 1), ncol = 3,
                  dimnames = list(0:n_cycles, v_names_states)))
names(m_trace) <- v_names_stages

for (stage in v_names_stages) {
  m_trace[[stage]][1, ] = v_init_states # subset matrix inside list
}

for (stage in v_names_stages) {
  for (t in 1:n_cycles) {
    m_trace[[stage]][t+1, ] <- m_trace[[stage]][t, ] %*% m_trans[[stage]]
  }
}


for (stage in v_names_stages) {
  m_trace[[stage]] <- as_tibble(m_trace[[stage]])
  m_trace[[stage]]["cycle"] <- 0:n_cycles
  m_trace[[stage]]["stage"] <- stage
}

df <- list_rbind(m_trace) |>
  pivot_longer(cols = c(Stable, Progressed, Dead), names_to = "state", values_to = "prob") |> 
  mutate(time = cycle / 50 * 3)


ggplot(data = df) +
  geom_line(aes(x = time, y = prob, colour = state, linetype = stage))

```





# Hesim

-   Not ideal because Delgado's model had no PSA.

## Model Structure

```{r}
strategies <- data.table(strategy_id = 1, 
                         strategy_name = "Usual care")
states <- data.table(state_id = 1:2,
                     state_name = )
```

## Patients

```{r}
patients <- data.table(patient_id = 1:4,
                       grp_id = 1:4,
                       stage = c("I", "II", "III", "IV"),
                       age = c(57.4, 62.4, 64.9, 68.1))

hesim_data <- hesim_data(strategies, patients, states)
labels <- get_labels(hesim_data)

```

## Outcomes

-   Using place-holder of 0.005 for background mortality.

-   Need to change to correct cycle length.

```{r}
median_pfs <- c(18.33, 6.25, 2.00, 1.60)
median_os <- c(19.50, 7.50, 3.20, 1.90)

tparams_def <- define_tparams({
  
  lambda_pfs <- - log(0.5) / median_pfs
  p_pp <- 1 - exp(-lambda_pfs * time)
  
  lambda_os <- - log(0.5) / (median_os - median_pfs)
  p_d <- 1 - exp(-lambda_os * time)
  
  # placeholder:
  p_mr <- 0.005
  
  list(
    tpmatrix = tpmatrix(
      C,   p_pp, p_mr,
      0,   C,    p_d + p_mr,
      0,   0,    1
    )
  )
}, times = 1:10)



```

```{r}
mod_def <- define_model(tparams_def = tparams_def)
```

