---
title: "Ovarian Cancer in Spain: Delgado 2019"
date: "`r conr::format_date()`"
author: "Connor Ballinger"
knit: conr::write_and_date
output: conr::format_html
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# library(here)
library(tidyverse)
library(readxl)
library(janitor)
library(knitr)
# library(hesim)
# library(data.table)

registerS3method("knit_print", "data.frame", conr::knit_print.data.frame)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```

# Overview

-   Epithelial OC (90% of OC is used).

-   3-state Markov (stable, progressed, death), stratified by the 4 stages. 

-   10-year horizon, societal perspective.

-   “The cycle length used was 3 weeks (21 days), which is the length of a chemotherapy cycle. Patients entered the model after they were diagnosed with EOC in the stable state... Those who progress remain in a post-progression state until they die.” 

-   Used Heintz 2006 for survival. Heintz only provided 5-year survival data – Delgado extrapolated based on exponential distribution ie time-constant transition probabilities.

-   Recall there may be no treatment effects applied - may just use Olaparib survival curves directly in an appropriate sub-population.

-   Note age-specific probabilities of death (incorporating OC mortality rate and background mortality rate).

    -   Add 2 rates and then calculate probability, or calculate probabilities and add? Delgado uses the latter. Always the latter.

# To do

-   Add yearly cohort with pop growth too. 

-   Add costs.

-   Construct a critique, including any Aus contextual factors.

-   Add equations.

# DARTH Framework (Alarid 2023)

-   Alarid-Escudero, F., Krijkamp, E., Enns, E. A., Yang, A., Hunink, M. G. M., Pechlivanoglou, P., & Jalal, H. (2023). A Tutorial on Time-Dependent Cohort State-Transition Models in R Using a Cost-Effectiveness Analysis Example. Medical Decision Making, 43(1), 21-41. doi:10.1177/0272989x221121747

-   Base R used in the interests of transparency and reducing the chance of running code without understanding it.

-   Objects are named with a prefix to indicate their type:

      -   n: number
      -   v: vector
      -   m: matrix
      -   a: array (generalisation of matrix)
      -   df: dataframe


-   Arrays are preferred to dataframes due to easier indexing and multiplication.

## Structure

```{r}
n_cycle_length <- 3 / 52 # 3 weeks
n_cycles <- round(10 / n_cycle_length) # 10 years

v_names_states <- c("Stable", "Progressed", "Dead")
v_names_stages <- c("I", "II", "III", "IV")
```

## Patients

```{r}
df_ages <- data.frame(cycle = 0:n_cycles, 
                      I = 57.4, 
                      II = 62.4, 
                      III = 64.9, 
                      IV = 68.1)
df_ages <- df_ages |>
  mutate(across(I:IV, ~ round(cycle * n_cycle_length + .x)))

df_ages_long <- df_ages |>
  pivot_longer(cols = -cycle, names_to = "stage", 
               values_to = "age", cols_vary = "slowest")
df_ages_long
```

## Transitions

-   Formulas in Delgado suggest probabilities vary over time, but they have assumed exponential so I don't see how.

-   Annual rates converted to 3-weekly rates by multiplying by 3/52.

### Background Mortality Rate

```{r}
df_life_table <- readxl::read_xlsx("abs_life_tables_2019-2021.xlsx", 
                                sheet = "Table_1.9",
                                skip = 6)
df_life_table <- df_life_table |> 
  select(Age, rate...7) |> 
  rename(age = Age, rate_annual_mort = rate...7)

# convert annual rate to 3-weekly rate

df_life_table <- df_life_table |> 
  mutate(age = as.integer(age),
         rate_bgr_mort = rate_annual_mort * n_cycle_length) |> 
  filter(between(age, 
                 min(df_ages_long$age), 
                 max(df_ages_long$age))) |> 
  select(age, rate_bgr_mort)

# ignore NA warning, that is due to ABS adding text below data

df_bgr_mort <- df_ages_long |> 
  left_join(df_life_table, by = "age")
df_bgr_mort
```

### Transition Rates - Heintz 2006

-   Median 5-year PFS and OS were published by stage. Delgado extrapolated assuming an exponential distribution.

$$S(t_{med}) = \frac{1}{2}$$
$$e^{-\lambda t_{med}} = \frac{1}{2}$$
$$t_{med} = \frac{\ln(2)}{\lambda}$$
-   Alternately:

$$\lambda = \frac{\ln(2)}{t_{med}}$$

```{r}
v_median_pfs <- c(18.33, 6.25, 2.00, 1.60)
v_median_os <- c(19.50, 7.50, 3.20, 1.90)

v_lambda_pfs <- - log(0.5) / v_median_pfs
v_3r_pfs <- v_lambda_pfs * n_cycle_length # 3-weekly rate

v_lambda_os <- - log(0.5) / (v_median_os - v_median_pfs)
v_3r_os <- v_lambda_os * n_cycle_length # 3-weekly rate

df_rates <- tibble(stage = v_names_stages,
                   rate_pfs = v_3r_pfs,
                   rate_os = v_3r_os)
df_rates
```

### All Rates to Probabilities

-   Convert rates to probabilities: $P = 1 - e^{-\lambda}$

```{r}
df_prob <- df_bgr_mort |> 
  left_join(df_rates, by = "stage") |> 
  mutate(across(.cols = starts_with("rate"),
                .fns = ~ 1 - exp(-.x),
                .names = "prob_{.col}")) |> 
  rename_with(.fn = ~ sub("_rate", "", .x), .cols = everything()) |> 
  mutate(prob_s_p = prob_pfs,
         prob_s_d = prob_bgr_mort,
         prob_p_d = prob_os + prob_bgr_mort) |> 
  select(cycle, stage, prob_s_p, prob_s_d, prob_p_d)
df_prob
```

-   Convert to array.

```{r}
# Split df_prob by Stage

df_prob_stages <- split(df_prob, df_prob$stage)
names(df_prob_stages) <- v_names_stages

# Initialise probability array

a_prob <- array(data = as.double(NA), 
                dim = c(3, 3, 4, (n_cycles+1)), 
                dimnames = list("from" = v_names_states, 
                                "to" = v_names_states,
                                "stage" = v_names_stages,
                                "cycle" = 0:n_cycles))
# Prevent some transitions

a_prob["Dead", "Stable", , ] <- 
  a_prob["Dead", "Progressed", , ] <- 
  a_prob["Progressed", "Stable", , ] <- 0

# Dead stay dead

a_prob["Dead", "Dead", , ] <- 1

# Use purrr::walk fn such that assignments made for each Stage

# S to P
v_names_stages |> 
  walk(\(x) a_prob["Stable", "Progressed", x, ] <<- 
         df_prob_stages[[x]]$prob_s_p)

# S to D
v_names_stages |> 
  walk(\(x) a_prob["Stable", "Dead", x, ] <<-  
         df_prob_stages[[x]]$prob_s_d)

# P to D
v_names_stages |> 
  walk(\(x) a_prob["Progressed", "Dead", x, ] <<- 
         df_prob_stages[[x]]$prob_p_d)

# S to S means neither S to P nor S to D occur
v_names_stages |> 
  walk(\(x) a_prob["Stable", "Stable", x, ] <<- 
         (1 - a_prob["Stable", "Progressed", x, ] - 
            a_prob["Stable", "Dead", x, ]))

# P to P means P to D does not occur
v_names_stages |> 
  walk(\(x) a_prob["Progressed", "Progressed", x, ] <<- 
         (1 - a_prob["Progressed", "Dead", x, ]))

a_prob[, ,c("I", "IV") , c(1, n_cycles + 1)] |> round(3)
```

## Checks

```{r}
sum(a_prob >= 0 & a_prob <= 1) - length(a_prob)

tracker <- 0

for (stage in v_names_stages) {
  for (t in 1:n_cycles+1) {
    tracker = tracker + rowSums(a_prob[, , stage, t]) - 1
  }
}
tracker # should be zero
```

## Construct Trace

```{r}
v_init_stages <- c(1155, 195, 1116, 681) # Delgado figures
# v_init_stages <- c(0.37, 0.06, 0.35, 0.22) # rough portions
# v_init_stages <- rep.int(1, times = 4)
v_init_states <- c(1, 0, 0) # all patients start in stable state

a_trace <- array(data = NA, 
                 dim = c(3, 4, n_cycles+1), 
                 dimnames = list("state" = v_names_states,
                                 "stage" = v_names_stages,
                                 "cycle" = 0:n_cycles))

a_trace[, , "0"] <- v_init_states %o% v_init_stages


for (stage in v_names_stages) {
  for (t in 2:(n_cycles + 1)) { # indexing by integer, not cycle name. cycle_name = integer - 1.
    a_trace[, stage, t] <- a_trace[, stage, t-1] %*% a_prob[, , stage, t]
  }
}

a_trace[, , c(1, n_cycles + 1)] |> 
  round()
```

## Gather Results

```{r}
df_trace <- as.data.frame(aperm(a_trace)) |>
  mutate(Cycle = 0:n_cycles) |> 
  pivot_longer(cols = -(Cycle), 
               names_to = c("Stage", "State"),
               names_sep = "\\.",
               values_to = "N") |> 
  mutate(Year = Cycle * n_cycle_length,
         State = factor(State, 
                        levels = c("Stable", "Progressed", "Dead"), 
                        ordered = TRUE))

ggplot(data = df_trace) +
  geom_line(aes(x = Year, y = N, colour = Stage)) +
  facet_wrap(vars(State)) #, scales = "free_y")

ggplot(data = df_trace) +
  geom_line(aes(x = Year, y = N, colour = State)) +
  facet_wrap(vars(Stage), nrow = 1)
```

## Costs

### Overarching figures:

```{r}
n_cpi_med <- 0.0077
n_cpi <- 0.0197
n_discount_rate <- 0.03
n_econ_growth <- 0.01


m <- matrix(rep.int(1, 20), nrow = 5) # for checking calcs
```

### Patients:

```{r}
# n_weight <- 66
# n_height <- 160
```

### Service-use rates:

```{r}
# n_brca_ref_rate <- 0.2
# n_brca_ref_visits <- 2
# n_brca_pos_rate <- 0.05
# n_brca_family_refs <- 5
# 
# v_hospital_6m <- c(1.2, 1.2, 2.1, 2.1)
# v_hospital_3w <- v_hospital_6m * 2 * n_cycle_length
# v_portion_hospitalised_6m <- c(0.154, 0.154, 0.482, 0.482) # what use is this?
# 
# v_emergencies_6m <- c(1.5, 1.5, 1.7, 1.7)
# v_emergencies_3w <- v_emergencies_6m * 2 * n_cycle_length
# v_portion_emergency_6m <- c(0.231, 0.231, 0.222, 0.222)
# 
# # note some received no treatment
# 
# v_tmt_surgery <- c(0.667, 0.198, 0.113, 0.088)
# v_tmt_neoadj_surgery <- c(0, 0, 0.144, 0.242)
# v_tmt_surgery_adjuv <- c(0.333, 0.802, 0.711, 0.582)
# 
# v_laparotomy <- c(1, 1, 1, 1)
# v_omentectomy <- c(0, 0.638, 1, 1)
# v_hysterectomy <- c(1, 1, 1, 1)
# v_bso <- c(0, 1, 1, 1)
# v_lymph <- c(0, 0, 0.75, 1)



df_key_params <- read_xlsx("delgado_parameters.xlsx", sheet = 7) |> 
  clean_names() |> 
  rename(param = x1) |> 
  filter(if_any(stage_i:stage_iv, ~ !is.na(.x)))
df_key_params <- df_key_params[-1, ] |> 
  mutate(across(stage_i:stage_iv, as.numeric))
m_params <- df_key_params |> 
  column_to_rownames(var = "param") |> 
  as.matrix()
m_params <- clean_names(m_params)
dimnames(m_params)
m_params
```

### Calculate volume of services

-   Should I name all these vectors using stage names, where relevant? eg v_init_stages.

```{r}
# initial surgery

v_init_stages
names(v_init_stages) <- v_names_stages

v_names_tmt_strategy <- dimnames(m_params)[[1]][10:13]
(v_names_surgery <- dimnames(m_params)[[1]][14:18])

(v_any_surgery <- (v_tmt_surgery + v_tmt_neoadj_surgery + v_tmt_surgery_adjuv) *
    v_init_stages)


(m_vol_surgery <- m_params[v_names_surgery, ] * 
    matrix(rep(v_any_surgery, 5), ncol = 4, byrow = TRUE))
```

### Service costs

```{r}
df_service_costs <- read_xlsx("delgado_parameters.xlsx", sheet = 1) |> 
  clean_names()
df_service_costs <- df_service_costs |> 
  select(-(references)) |> 
  rename(service = x1,
         cost = costs_2016) |> 
  drop_na(cost) |> 
  mutate(service = make_clean_names(service),
         cost = as.double(cost))
df_service_costs
v_cost_services <- deframe(df_service_costs)
names(v_cost_services)


```

### Total Cost of Services

```{r}
(m_surg_cost <- m_vol_surgery *
  v_cost_services[c("laparotomy", 
                    "abdominal_total_hysterectomy", 
                    "lymphadenectomy",
                    "omentectomy",
                    "bilateral_salpingoophorectomy")])
```

### Drug costs

```{r}
df_drug_costs <- read_xlsx("delgado_parameters.xlsx", sheet = 2, skip = 1) |> 
  clean_names()
df_drug_costs <- df_drug_costs |> 
  select(-(list_price_lp)) |> 
  rename(drug = x1, 
         cost = lp_rdl_vat) |> 
  drop_na(cost)
df_drug_costs
```

### Drug doses

```{r}
df_drug_doses <- read_xlsx("delgado_parameters.xlsx", sheet = 3, skip = 2) |> 
  clean_names()
df_drug_doses <- df_drug_doses |> 
  rename(drug = x1,
         time_hours = x6)
df_drug_doses
```

### Drug rates

```{r}
df_drug_use <- read_xlsx("delgado_parameters.xlsx", sheet = 4, skip = 2) |> 
  clean_names()
df_drug_use <- df_drug_use |> 
  rename(drug = x1,
         line1_stages_i_ii = stages_i_and_ii,
         line1_stages_iii_iv = stages_iii_and_iv,
         line2 = x4,
         line3 = x5,
         line4 = x6,
         line5 = x7,
         line6 = x8,
         line7 = x9)
df_drug_use
```

### Palliative drugs

```{r}
df_drug_pall <- read_xlsx("delgado_parameters.xlsx", sheet = 5, skip = 2) |> 
  clean_names()
df_drug_pall <- df_drug_pall |> 
  rename(drug = x1,
         use_pre_terminal = use_of_drugs_2,
         daily_dose_pre_terminal = daily_dose_mg_m_l_3,
         use_post_terminal = use_of_drugs_4,
         daily_dose_post_terminal = daily_dose_mg_m_l_5)
df_drug_pall
```

